Diagnoses and cell types

### 2025-10-08 analyses/infercnv-consensus-cell-type/exploratory-notebooks/reference-query-counts.Rmd
ensure we are only using most up to date consensus cells with scimilarity, meaning no assessment of multiplexed

library_diagnosis_df <- all_library_diagnosis_df |>
  # don't consider cell line (or multiplexed?)
  # for now at least, we are retaining multiplexed
  dplyr::filter(!is_cell_line)
```

@@

      # bump guess_max to kill problems() warnings
      readr::read_tsv(file, guess_max = 15000) |>
       dplyr::select(
         scpca_project_id = project_id,
         scpca_library_id = library_id,
         barcodes,
         consensus_annotation,
         consensus_ontology
       )
  }) |>
  dplyr::bind_rows()
```

@@

We should also check the eligible `SCPCP000009` libraries to see if they are multiplexed:

```{r}
library_diagnosis_df |>
  dplyr::filter(scpca_project_id == "SCPCP000009", scpca_library_id %in% usable_libraries_df$scpca_library_id) |>
  dplyr::count(is_multiplexed)
```

@@

We are able to run about 51% of libraries through `inferCNV`, based on consensus cell types derived from `SingleR`, `CellAssign`, and `SCimilarity` labels.
Most broad diagnosis groups are represented, except `Mixed phenotype acute leukemia` (n=18) and a handful of other solid tumors (n=3 among Gastrointestinal stromal tumor and Hepatoblastoma).
All eligible projects are represented as well.

Note that we are including multiplexed libraries from project `SCPCP000009` here as well; 12 libraries from that project from are eligible, and all but of of those are multiplexed.
These libraries would effectively be using pooled references in `inferCNV`, since we did not demultiplex libraries.

>>>>

library_diagnosis_df <- all_library_diagnosis_df |>
  # don't consider cell line or multiplexed in this notebook
  # the multiplexed samples were not processed with SCimilarity in OpenScPCA-nf
  dplyr::filter(!is_cell_line, !is_multiplexed)
```

@@

      # bump guess_max to kill problems() warnings
      consensus_df <- readr::read_tsv(file, guess_max = 15000)

      # ensure we have the most up-to-date files - aka, with SCimilarity
      stopifnot("Missing SCimilarity in the consensus files" = "scimilarity_celltype_annotation" %in% colnames(consensus_df))

      consensus_df |>
        dplyr::select(
         scpca_project_id = project_id,
         scpca_library_id = library_id,
         barcodes,
         consensus_annotation,
         consensus_ontology
       )
  }) |>
  dplyr::bind_rows()
```

@@

@@

We are able to run about 50\% of libraries through `inferCNV`, based on consensus cell types derived from `SingleR`, `CellAssign`, and `SCimilarity` labels.
Most broad diagnosis groups are represented, except `Mixed phenotype acute leukemia` (n=18) and a handful of other solid tumors (n=3 among Gastrointestinal stromal tumor and Hepatoblastoma).
All eligible projects are represented as well.

Note that it was not possible to fully assess multiplexed libraries here since they did not undergo `SCimilarity` (and hence their consensus cell types are not updated) inference in `OpenScPCA-nf`. 
They will be included when running through `scpca-nf`, however, assuming they have sufficient reference cells; it is important to bear in mind that these libraries would effectively be using pooled references in `inferCNV`, since we did not demultiplex libraries.
